\textbf{Analyzing Machupo virus-receptor binding by molecular dynamics simulations}

### 2014-01-06 paper.tex
Last update before sending to Sara and Andy.

Here we propose that much of these complexities can be avoided if all we are interested in is a relative comparison of the effects of different mutations on protein-protein interactions, rather than measuring an absolute or relative binding affinity with experimentally realistic units. We impart a pulling force within an all-atom molecular dynamics simulation on one member of the complex while the other is held in place. Then, we measure the force required for dissociation \citep{Gumbart2012,Lu1999,Park2004,Is2001A,Is2001B}. Although such biasing techniques are commonly used in protein-ligand binding problems, they are less commonly applied to protein--protein interactions, and almost never to mutational analysis in a protein--protein system. This is largely the result of free energy convergence difficulties and computational limitations \citep{Cuendet2008,Cuendet2011}. Using a proxy for relative binding affinity rather than caluclating absolute affinities can solve these problems. Here, as proxies, we use the maximum applied force required for separation and the area under the force-versus-distance curve (AUC). For comparison, we also calculate relative free energy differences using the traditional dual topology FEP paradigm, and we show that the two approaches yield congruent results.

@@

We used the final state from each equilibrated system to restart another MD simulation. Our steering protocol is fundamentally similar to \citet{Cuendet2008} with slightly different parameter choices. Perhaps the one significant difference lies in our choosing to not use a thermostat or barostat. We can make this choice because we are not trying to calculate the binding free energy by any physically rigorous approach (the Jarzynski inequality being one example). Following equilibration, the final state of each simulation was used to generate a configuration file fixing the $\alpha$-carbon on residues 1, 58, 73-83, 96, 136, 137, 138, and 161 (again with linear numbering) in the hTfR1. These residues were selected as they are far from the binding interface and sufficiently distributed to prevent any orientational motion of the receptor relative to the viral spike protein. The center of mass of the $\alpha$-carbons of all residues (163-318 in linear numbering) in GP1 received an applied force during the simulation. The NAMD convention does not actually apply a force to all $\alpha$-carbon atoms but rather uses the selection to compute an initial center of mass. Then, during the steering run, the single center of mass point is pulled with the parameters desribed below. We used the same force field parameters (exclude, cutoff, switching, etc.), the same integrator parameters (time step, rigidbonds on, all molecular being wrapped, etc.), and the same particle mesh ewald parameters as in equilibration. Periodic boundary conditions were incorporated as part of the system (as is the convention in NAMD restart) and PME was again used to approximate full system electrostatics.

@@

The statistical package R was used for all further analysis and visualization. Each of the 50 independent trajectories per mutant produced a fairly noisy force curve. The force curves for each mutant were smoothed by using the smooth.spline() and predict() functions in R with default settings over all replicates as in Figure~\ref{fig:force_curve}. The two primary descriptive statistics we used were maximum interpolated applied force and total area under the interpolated curve (AUC). We tested for signifiant differences in maximum force or AUC by carrying out t tests for all pairwise combinations (each mutant compared to each other mutant), using the pairwise.t.test() function in R. We adjusted p values to correct for multiple testing using the False-Discovery-Rate (FDR) method \citep{BenjaminiHochberg1995}. The ggplot \citep{ggplot} package was used to generate most of the figures.

@@

We analyzed the GP1/hTfR1 system using two molecular dynamics techniques. First, by carrying out SMD using a known force constant and pulling with a constant velocity, we could calculate the applied force during protein--protein dissociation \citep{Cuendet2008,Cuendet2011}. A typical averaged force curve comparison can be seen in Figure~\ref{fig:force_curve}. As seen in Figure~\ref{fig:force_curve}, the dissociation distance was relatively consistent among mutants. {\color{red}\emph{I would still like to see an array of figures with all averaged force curves.}} The supplementary movie visually illustrates the separation distance between peptide domains. The quantities maximum applied force and AUC were derived from the force-versus-distances curves. Their summary statistics are reported in Table~\ref{tab:summary_statistics}. As we are more interested in the phenotypic impact of interface mutations we avoided many of the more physically rigorous, but technically complicated calculations that are possible with SMD \citep{Is2001A,Is2001B}.

@@

\section{Acknowledgements}
This work was carried out using high-performance computing resources provided by the High Performance Computing Center (HPCC) at Texas Tech University at Lubbock \\ (http://www.hpcc.ttu.edu) and the Texas Advanced Computing Center (TACC) at The University of Texas at Austin (http://www.tacc.utexas.edu). We would like to thank Bryan Sutton for opening access to the Hrothgar cluster and the reviewers Ilan Samish and Matteo Masetti for their helpful comments on this work.

@@

\clearpage
\begin{figure}[p]
\centerline{\includegraphics[width=6in]{figures/rmsfs.pdf}}
\caption{\label{fig:rmsf}\title{RMSF values during equilibration.} The RMSF values for every site in the bound complex computed during the equilibration phase of the protocol. Each color represents the average over 20 trajectories of a single mutant. Indices 17-25 are the hTfR flexible loop. The plot shows the flexibility of each site is essentially independent of mutation, and two sites (indices 17 and 18) above 0.72~\AA\ are a part of the flexible loop in the free receptor. However, these two residues are not actually found in the protein--protein interface, but rather are almost completely solvent exposed with the virus bound. {\color{red}\emph{Please highlight the flexible loop, as in Fig. 4.}}}
\end{figure}

>>>>

Here we propose that much of these complexities can be avoided if all we are interested in is a relative comparison of the effects of different mutations on protein-protein interactions, rather than measuring an absolute or relative binding affinity with experimentally realistic units. We impart a pulling force within an all-atom molecular dynamics simulation on one member of the complex while the other is held in place. Then, we measure the force required for dissociation \citep{Lu1999,Is2001A,Is2001B,Park2004,Gumbart2012,Minoetal2013}. Although such biasing techniques are commonly used in protein-ligand binding problems, they are less commonly applied to protein--protein interactions, and almost never to mutational analysis in a protein--protein system. This is largely the result of free energy convergence difficulties and computational limitations \citep{Cuendet2008,Cuendet2011}. Using a proxy for relative binding affinity rather than caluclating absolute affinities can solve these problems. Here, as proxies, we use the maximum applied force required for separation and the area under the force-versus-distance curve (AUC). For comparison, we also calculate relative free energy differences using the traditional dual topology FEP paradigm, and we show that the two approaches yield congruent results.

@@

We used the final state from each equilibrated system to restart another MD simulation. Our steering protocol is fundamentally similar to \citet{Cuendet2008} with slightly different parameter choices. Perhaps the one significant difference lies in our choosing to not use a thermostat or barostat. We can make this choice because we are not trying to calculate the binding free energy by any physically rigorous approach (the Jarzynski inequality being one example). Following equilibration, the final state of each simulation was used to generate a configuration file fixing the $\alpha$-carbon on residues 1, 58, 73-83, 96, 136, 137, 138, and 161 (again with linear numbering) in the hTfR1. These residues were selected as they are far from the binding interface and sufficiently distributed to prevent any orientational motion of the receptor relative to the viral spike protein. The center of mass of the $\alpha$-carbons of all residues (163-318 in linear numbering) in GP1 received an applied force during the simulation. The NAMD convention does not actually apply a force to all $\alpha$-carbon atoms but rather uses the selection to compute an initial center of mass. Then, during the steering run, the single center of mass point is pulled with the parameters described below. We used the same force field parameters (exclude, cutoff, switching, etc.), the same integrator parameters (time step, rigidbonds on, all molecular being wrapped, etc.), and the same particle mesh ewald parameters as in equilibration. Periodic boundary conditions were incorporated as part of the system (as is the convention in NAMD restart) and PME was again used to approximate full system electrostatics.

@@

The statistical package R was used for all further analysis and visualization. Each of the 50 independent trajectories per mutant produced a fairly noisy force curve. The force curves for each mutant were smoothed over all replicates by using the smooth.spline() and predict() functions in R with default settings. The two primary descriptive statistics we used were maximum interpolated applied force and total area under the interpolated curve (AUC). We tested for signifiant differences in maximum force or AUC by carrying out t tests for all pairwise combinations (each mutant compared to each other mutant), using the pairwise.t.test() function in R. We adjusted p values to correct for multiple testing using the False-Discovery-Rate (FDR) method \citep{BenjaminiHochberg1995}. The ggplot \citep{ggplot} package was used to generate most of the figures.

@@

We analyzed the GP1/hTfR1 system using two molecular dynamics techniques. First, by carrying out SMD using a known force constant and pulling with a constant velocity, we could calculate the applied force during protein--protein dissociation \citep{Cuendet2008,Cuendet2011}. A typical averaged force curve comparison can be seen in Figure~\ref{fig:force_curve}, and individual images of all averaged force curves are available in the associated github repository, in folder figures/force\_curves. As seen in Figure~\ref{fig:force_curve}, the dissociation distance was relatively consistent among mutants.  The supplementary movie visually illustrates the separation distance between peptide domains. The quantities maximum applied force and AUC were derived from the force-versus-distances curves. Their summary statistics are reported in Table~\ref{tab:summary_statistics}. As we are more interested in the phenotypic impact of interface mutations we avoided many of the more physically rigorous, but technically complicated calculations that are possible with SMD \citep{Is2001A,Is2001B}.

@@

\section{Acknowledgements}
This work was carried out using high-performance computing resources provided by the High Performance Computing Center (HPCC) at Texas Tech University at Lubbock (http://www.hpcc.ttu.edu) and the Texas Advanced Computing Center (TACC) at The University of Texas at Austin (http://www.tacc.utexas.edu). We would like to thank Bryan Sutton for opening access to the Hrothgar cluster and the reviewers Ilan Samish and Matteo Masetti for their helpful comments on this work.

@@

\clearpage
\begin{figure}[p]
\centerline{\includegraphics[width=6in]{figures/rmsfs.pdf}}
\caption{\label{fig:rmsf}\title{RMSF values during equilibration.} The RMSF values for every site in the bound complex computed during the equilibration phase of the protocol. Each color represents the average over 20 trajectories of a single mutant. Indices 17-25 are the hTfR flexible loop. The plot shows the flexibility of each site is essentially independent of mutation, and two sites (indices 17 and 18) above 0.72~\AA\ are a part of the flexible loop in the free receptor. However, these two residues are not actually found in the protein--protein interface, but rather are almost completely solvent exposed with the virus bound.}
\end{figure}
